---
title: Showcasing how R and Python can talk!
author: ''
date: '2020-12-10'
slug: showcasing-how-r-and-python-can-talk
categories: []
tags: []
description: ''
---

In more ways than one, Python is a very powerful programming language. Thus far, however, all of my projects have been conducted using R. This write-up is a demonstration of how Python can accomplish many of the same things, and interestingly, interact with R in the same environment (using RStudio).

To start, let's take a look at the `quakes` dataset in R. The `quakes` dataset contains the locations of 1000 earthquakes near Fiji since 1964.
```{r}
#in R
library(tidyverse)
#saving quakes dataset as quakesR
quakesR <- quakes
#taking a peek at quakes
glimpse(quakesR)
```
Using R, it's pretty easy to find out manipulate this data and find out about it.
```{r}
#finding summary statistics for depth/mag/stations in R
quakesR %>% 
  summarize_at(vars(depth, mag, stations),
               funs(mean, sd))
```

What if we're interested in seeing what the distribution of `depth` looks like?
```{r}
#in R
#making a histogram
hist(quakes$depth, main = "Histogram of Depth", xlab = "Depth")
```

How about if we want to see a plot of the quakes' geographic locations?
```{r fig.width=4, fig.height=4}
#in R
#making scatterplot
plot(quakes$long, quakes$lat, xlab = "Longitude", ylab = "Latitude", main = "Latitude vs Longitude")
```

R is pretty cool! But what if we wanted to do all of this in Python instead? Instead of using Python to read in a new dataset, we can actually just use the dataset we've already brought into the global environment using a package called "reticulate."
```{r}
#in R
#let's go ahead and load up reticulate
library(reticulate)
```

Now we can use the R dataset (and for that matter, any variable/value created in R) in Python! To use objects from R, all we have to do is put "r." before the object in a Python code chunk
```{python}
#in Python
#let's take a peek at quakesR
r.quakesR.head()
```

Awesome! Now what if we wanted to find the mean/sd of `depth`, `mag`, and `stations` using Python like we did above?
```{python}
#in Python
#importing library
import pandas as pd

#describe from the pandas library actually provides some additional summary stats as well!
r.quakesR[['mag', 'depth', 'stations']].describe()
```

Pretty cool, these are the same results as in R.

What about a histogram of `depth`?
```{python}
#in Python
#importing library
import matplotlib.pyplot as plt

#making histogram of depth
plt.hist(r.quakesR['depth'])
plt.xlabel('Depth')
plt.ylabel('Frequency')
plt.show()
```

What about if we want a plot of the quakes' geographic locations using Python? Here's a pretty elaborate version!
```{python}
#in Python
side, main = plt.subplots()
fiji = plt.imread("http://valorielord.com/default/cache/file/D4256462-E44C-32F4-62C9A3C4AA762918_bodyimage.png")
main.imshow(fiji, extent = [135,215,-40,0])
#note that to use the quakesR dataset, we have to include "r."
plt1 = main.scatter(x = r.quakesR['long'], y = r.quakesR['lat'], alpha = .2, c = r.quakesR['depth'])
side.colorbar(plt1)
plt.xlabel("Longitude")
plt.ylabel("Latitude")
plt.title("Latitude vs Longitude")

plt.show()

```

To go a little farther using Python, what if we were interested in the mean magnitude of earthquakes above and below the median depth of all earthquakes?
```{python}
#importing Python
#importing library
import numpy as np

#lets create a variable called meddep (median depth)
meddep = np.median(r.quakesR.depth)

#the median depth is 247
print(meddep)

#finding mean of magnitude when depth is greater than meddep (deeper quakes)
r.quakesR[r.quakesR.depth > meddep].mag.mean()

#finding mean of magnitude when depth is less than meddep (shallower quakes)
r.quakesR[r.quakesR.depth < meddep].mag.mean()
```

These results show us that when depth is greater than the median depth, the mean magnitude is less than that of when depth is less than the median depth!

But what if we're interested in finding this out using R? We don't have to make a whole new `meddep` variable, Python and R can talk back and forth! To access variables that Python created in an R code chunk, all we need is "py$"
```{r}
#in R
#calling on meddep from Python
py$meddep

#finding mean of magnitude when depth is greater than meddep (deeper quakes)
mean(quakesR[quakesR$depth > py$meddep,]$mag)

#finding mean of magnitude when depth is less than meddep (shallower quakes)
mean(quakesR[quakesR$depth < py$meddep,]$mag)
```

These results are one and the same! That's all for this short demo, hope it was helpful in showing how R and Python can talk with each other in RStudio!

![To wrap up, here's a pretty picture I took in Sydney!](/img/IMG_9692.HEIC){width=100%}



